Our Galactic Midplane Is Probably Wrong, and Not by a Little

### 2018-12-20 paper/ms.tex
Add to discussion section

\section{Discussion} \label{sec:discussion}
{\em Gaia} DR2 has ushered in a new era of Galactic dynamics. High-precision
phase space measurements have been made for an unprecedented number of stars.
In the face of such large data, it is worth considering the goal of such
measurements --- a goal which should inform the usage of such data.

In this work we considered one such question: given the current position in
phase space of a single star, what can we infer about that star's past phase
space trajectory? Current methodology involves using a smoothed, analytic form
for the Galactic potential (with perhaps crude implementations of spiral arms
or sattelites), fitting such form to some selection of observations, and
integrating from the currently measured position in phase space. This orbit
integration is also commonly accompanied with a transformation into
action-angle space. Other than the integration, every step of that process
should be viewed with significant scrutiny.

The more general problem of measuring the orbits of stars in cosmological
simulations is essentially unexplored in this work and the literature, and we
delay discussion to future work.

>>>>

\section{Discussion} \label{sec:discussion}
We have shown that, in simulation, the ``local midplane'' can have strong
azimuthal dependence. The implications of this finding are not clear, but
clearly demonstrate that assuming our local midplane is the same as the global
midplane is insufficient for extracting the most dynamical information from
the {\em Gaia} mission.

A possible interpretation of our finding is that one should use a star's local
midplane in order to compute dynamical quantities like the vertical action.
Under the assumption that this interpretation is correct, we now consider the
implications for dynamical studies of disrupting open clusters (the original
motivation of this study).

Suppose an open cluster of mass $M_c$ on an orbit with actions $\bm{J}_c$ is
disrupted, and that the cluster is currently at an orbit phase of
$\bm{\theta}_0$. We would like to determine all stars in the {\em Gaia}
dataset\footnote{This is likely to be more successful with DR3/4 than DR2,
since upcoming data releases will contain a far greater number of stars with
radial velocities --- though targeted, follow-up radial velocity surveys may
allow this program to be performed sooner.} that are likely to be ejected
members from that cluster. One could proceed by integrating the cluster's
orbit and selecting all stars within a certain distance of that orbit,
accounting for the fact that the stream will not exactly follow the orbit of
the cluster \citep[e.g.][]{2011MNRAS.413.1852E}.

One could then make a further selection by requiring that the action of each
field star $i$ be close to the cluster's actions to within some bound, i.e.
$\bm{J}_i - \bm{J}_c \in V_e$ where $V_e$ is some volume in action space
enclosing a region of acceptable error. A natural choice for $V_e$ is to
center it at the origin and allow its extent to be the intrinsic action-space
dispersion of a disrupted cluster. This dispersion can be estimated as
\citep[\S~8.3.3][]{2008gady.book.....B}
\beq \label{eq:action_disp}
\frac{\Delta J_i}{J_i} \sim \left(\frac{m}{M}\right)^{1/3}\text{,}
\eeq
where $m$ is the mass of the cluster and $M$ is the mass enclosed by the
cluster's orbit. A more sophisticated treatment of the action-space dispersion
of disrupted clusters can be obtained by following
\citet{2011MNRAS.413.1852E}, but such a treatment is premature for this work.

The act of comparing a field star's action to a cluster's action (i.e. in
computing $\bm{J}_i - \bm{J}_c$) is subject to significant error if the
cluster and field star have significantly different local midplanes and an
observer assumed a single global midplane.

To explore the magnitude of this effect, we consider the following program.
First, we assume a cluster mass $m$ and estimate the expected action-space
dispersion from Equation~\ref{eq:action_disp}. We then consult
Figure~\ref{fig:many_orbit_wrong_ref} to determine the $z$ offset necessary in
order for the induced error in the vertical action to be the same as the
expected action-space dispersion. If this were the case, an observer would
have to account for the different midplane of the field star and the cluster.
We repeat this analysis for each of the three orbits considered in
Section~\ref{ssec:quant}.

With a characteristic $z$ offset in hand, we consult Figure~\ref{fig:deltaphi}
to convert this $z$ offset into a $\Delta \phi$, which we convert to an
arclength or characteristic distance (the difference between the two will be
negligible for the small $\Delta \phi$'s we will compute). Therefore, if an
observer is attempting to reconstruct a cluster with a given mass on a given
orbit, we will give a characteristic distance between a field star and the
cluster within which an observer might not need to consider differences in
their respective local midplanes.

From the simulation we have the best handle on the $z$ offset
(Figure~\ref{fig:midplane}), and this $z$ offset has the greatest impact on
the vertical action $J_z$ (Figure~\ref{fig:many_orbit_wrong_ref}). Therefore,
we only consider the impact of a $z$ offset on the vertical action $J_z$.

% {\em Gaia} DR2 has ushered in a new era of Galactic dynamics. High-precision
% phase space measurements have been made for an unprecedented number of stars.
% In the face of such large data, it is worth considering the goal of such
% measurements --- a goal which should inform the usage of such data.

% In this work we considered one such question: given the current position in
% phase space of a single star, what can we infer about that star's past phase
% space trajectory? Current methodology involves using a smoothed, analytic form
% for the Galactic potential (with perhaps crude implementations of spiral arms
% or sattelites), fitting such form to some selection of observations, and
% integrating from the currently measured position in phase space. This orbit
% integration is also commonly accompanied with a transformation into
% action-angle space. Other than the integration, every step of that process
% should be viewed with significant scrutiny.

% The more general problem of measuring the orbits of stars in cosmological
% simulations is essentially unexplored in this work and the literature, and we
% delay discussion to future work.
